---
title: "Module 3"
author: "Jason Andrews"
date: "14 FEB 23"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))

```{r, error=TRUE}
mycalc<- function(x,y,fun) {
    if(fun == "sum") {
    o <- x+y
  } else if(fun == "subtract") {
    o <- x-y
  } else if(fun == "divide") {
    o <- x/y
  } else if(fun == "multiply") {
    o <- x*y
  } else {
    o <- paste("Not a recognized function")
  }
  return(o)
}
s<-sample(1:10, 10, replace=TRUE)
s
t<-matrix(runif(30, min = 1, max = 100), ncol = 3)
t
mycalc(s,t,"multiply")
```
## 1. Create the following vectors:

## - An integer vector of length 20, to be named v1, using the sample function to randomly select 20 values from 0 through 100, use the option “replace = TRUE” (which means that you can the same value randomly selected more than one time)

## - A second integer vector (named v2), also of length 20, but selecting 20 values from the range 1:10.

## - A numeric vector of length 20 (named v3), using the runif function with the minimum set to 0 and the maximum to 1.

## - A character vector of length 20 (named v4), using randomly selected values from the built-in dataset LETTERS. You will use v2 to select the values out of LETTERS (this means that you are using v2 as an index).

## - In making v1, v2, v3, you will use the set.seed function so that you can get the exact same results each time you run your code to make them. Use a seed value of 1.
```{r, error=TRUE}
set.seed(1)
v1<-sample(1:100, 20, replace=TRUE)
v1
v2<-sample(1:10, 20,replace=TRUE)
v2
v3<-runif(20, min=0,max=1)
v3
v4<-sample(LETTERS, 20, replace=TRUE)
v4
```
## 2. Create a matrix m1 from v1 and v2, by binding the two vectors together as columns. Create another (m2) from v2 and v3, again by binding columns. Create a third (m3) by binding v3 and v4.
```{r, error=TRUE}
m1<-cbind(v1,v2)
m1
m2<-cbind(v2,v3)
m2
m3<-cbind(v3,v4)
m3
```
## 3. Create a data.frame (named DF) by bindings the columns of v1, v2, v3, and v4.
```{r}
DF <- data.frame(v1, v2, v3, v4)
DF
```
## - Question: why shouldn’t you use the name “df” to name this object?
Because "df" is a reserved word in R, and it would be confusing. 

## - Change the column names of DF to “a”, “b”, “c”, and “grp”.
```{r}
colnames(DF) <- c("a", "b", "c", "grp")
```

## - Question: why is it okay to rename a column in DF “c”, but not to rename v3 that name?
"c" is a common command (concatenate) in R, and naming a vector this would reduce readability. Also, column names are specific to data frames. 

## - Look at the str of DF. What kind of data type is “grp”? How would you change it between character and factor and back again?
```{r}
str(DF)
DF$grp <- as.factor(DF$grp)

DF$grp <- as.character(DF$grp)
```
## 4. Demonstrate the use of your new function:
## - Use mycalc to sum and divide the following pairs of objects:
## - Capture the output of each operation in objects d1 through d5 for the divisions, and s1 through s5 for the sums.
v1, v2
```{r}
d1<-mycalc(v1,v2,"divide")
s1<-mycalc(v1,v2,"sum")
```
v2, v3
```{r}
d2<- mycalc(v2,v3,"divide")
s2<- mycalc(v1,v2,"sum")
```
v1, v3
```{r}
d3<-mycalc(v1,v3,"divide")
s3<-mycalc(v1,v2,"sum")
```
m1, m1
```{r}
d4<-mycalc(m1,m1,"divide")
s4<-mycalc(v1,v2,"sum")
```
m1, m2
```{r}
d5<-mycalc(m1,m2,"divide")
s5<-mycalc(v1,v2,"sum")
```
## Using mycalc with DF:
## Multiply row1, column1 by row1, column2
```{r}
DF
mycalc(DF[1,1],DF[1,2],"multiply")
```
## Multiply rows 1-10, column 1-3 by rows1-10, columns 1-3
```{r}
mycalc(DF[(1:10),(1:3)],DF[(1:10),(1:3)],"multiply")
```

## Divide the values in columns “a” and “b” by the values in column “c” of DF corresponding to the rows in “grp” that have values “E”. In selecting the columns, use their names, rather than their index numbers, to select them
```{r}
mycalc(DF[,c("a","b")],DF[DF$grp == "F", "c"], "divide")
```
## - Repeat the last of the aforementioned calculations, but this time do not use mycalc. Simply do it using the two DF subsets with the / operator
```{r}
DF[,c("a","b")]/DF[DF$grp == "F", "c"]
```

## 5. Demonstrate the use of *apply:

## - Combine v1, m1, and m2 into a list l

```{r}
l<-list(v1,m1,m2)
l
```

## - Write an lapply that applies sum to l. This one does not need an anonymous function
```{r}
lapply(l,sum)
```

## - Now write an lapply that iterates over the elements of l, and uses your mycalc to multiply each element of l by v2
## - Hints:
## - You will need to use an anonymous function (function(x))
## - The interior of your function (i.e. the part between ({}) use the mycalc function.

```{r}
lapply(l, function(x) { 
  mycalc(x,v1,"multiply")
})

```
## - Do the same but with sapply
```{r}
sapply(l, function(x) { 
  mycalc(x,v1,"multiply")
})
```

