```{r}
a <- 1:10
b <- a
names(b) <- letters[1:10]
?names
b
```
Note that we use [] with letters, (). That’s because letters is a vector from which we are extracting a subset of values, as opposed to an object to which we are applying a function, in which case we would enclose the object with ()–as we did with names(a).
```{r}
b["a"]
#> a 
#> 1
b[c("d", "e")]
#> d e 
#> 4 5
b[c("a", "e", "j")]
#>  a  e  j 
#>  1  5 10
b["j"]
#>  j 
#> 10
b[-"a"]
#> Error in -"a": invalid argument to unary operator
b[-c("a", "c")]
#> Error in -c("a", "c"): invalid argument to unary operator
```
We can also index using logical operators to select elements of vectors based on their values.
```{r}
# #1
a[a > 5]
#> [1]  6  7  8  9 10
#
# #2
a[a >= 2 & a < 7]
#> [1] 2 3 4 5 6
#
# #3
a[a == 7 | a == 2]
#> [1] 2 7
#
# #4
a[a %in% c(1, 10)]
#> [1]  1 10
#
# #5
b[b %in% 2:3]
#> b c 
#> 2 3
```
Line #4: Select from a all values of a that occur within a vector containing 1 and 10 (this is the same as: select from a all values of a that equal 1 or equal 10)
Line #5: Select from b all values of b that occur within a vector containing 2 and 3

```{r}
a > 5
#>  [1] FALSE FALSE FALSE FALSE FALSE  TRUE  TRUE  TRUE  TRUE  TRUE
a >= 2 & a < 7
#>  [1] FALSE  TRUE  TRUE  TRUE  TRUE  TRUE FALSE FALSE FALSE FALSE
a == 7 | a == 2
#>  [1] FALSE  TRUE FALSE FALSE FALSE FALSE  TRUE FALSE FALSE FALSE
a %in% c(1, 10)
#>  [1]  TRUE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE  TRUE
b %in% 2:3
#>  [1] FALSE  TRUE  TRUE FALSE FALSE FALSE FALSE FALSE FALSE FALSE
```

```{r}
which(a > 5)
#> [1]  6  7  8  9 10
which(b %in% 2:3)
#> [1] 2 3
```


```{r}
l <- list(1, 1:10, c("a", "b", "c", "d"))  # unnamed list
l2 <- l
names(l2) <- letters[1:3]  # named list
l 
#> [[1]]
#> [1] 1
#> 
#> [[2]]
#>  [1]  1  2  3  4  5  6  7  8  9 10
#> 
#> [[3]]
#> [1] "a" "b" "c" "d"
l2
#> $a
#> [1] 1
#> 
#> $b
#>  [1]  1  2  3  4  5  6  7  8  9 10
#> 
#> $c
#> [1] "a" "b" "c" "d"
```


```{r}
# Chunk 3
names(l2$c) <- letters[1:4]
#
# #1
l[[2]][2:3]
#> [1] 2 3
#
# #2
l2$b[2:3]
#> [1] 2 3
#
# #3
l[2][[1]][2:3]
#> [1] 2 3
#
# #4
l2["b"][[1]][2:3]
#> [1] 2 3
#
# #5
l[[3]][c(1, 4)]
#> [1] "a" "d"
#
# #6
l2$c[c(1, 4)]
#>   a   d 
#> "a" "d"
#
# #7
l2$c[c("a", "d")]
#>   a   d 
#> "a" "d"
#
# #8
l2["c"][["c"]][c("a", "d")]
#>   a   d 
#> "a" "d"
#
# #9
l[2:3][3]  # doesn't work
#> [[1]]
#> NULL
```


```{r}
set.seed(1)
a <- sample(0:100, size = 10, replace = TRUE)
a
names(a) <- letters[1:10]
b <- a  # copy we will modify
l <- list(e = 1, f = 1:10, g = a)
l2 <- l  # copy we will modify
l
```
```{r}
b[1] <- -99 
b
b["j"] <- "z" 
b
b[c("b", "f")] <- 9999
b
b[3:4] <- c(-1, -2)
b
b[5:length(b)] <- 10000:10001 
b
b[length(b) - 1] <- 0:10  
b
```
```{r}
l2
l2[[1]] <- c(1, 4)
l2[[1]]

l2$f[c(1, 10)] <- c(-1, 1000)
l$f

l2$f

l2[[3]][letters[1:4]] <- 1:4

l[[3]]

l2[[3]]

l2$myfun <- function(x) x * 10  
l
l2

l2$myfun(l2$f)
```

2.1.4 PRACTICE
In 2.1.1, what class of object is a? Recreate a in your own script and apply a function to it to get the answer.
```{r}
a <- 1:10
typeof(a)
```
In 2.1.2, Chunk 3 #3, #4, and #8, why do we have to use three sets of brackets to get access to the vector elements? Hint: pay attention to the first set of brackets.

```{r}
# Chunk 3
names(l2$c) <- letters[1:4]
#
# #1
l[[2]][2:3]
#> [1] 2 3
#
# #2
l2$b[2:3]
#> [1] 2 3
#
# #3
l[2][[1]][2:3]
#> [1] 2 3
#
# #4
l2["b"][[1]][2:3]
#> [1] 2 3
#
# #5
l[[3]][c(1, 4)]
#> [1] "a" "d"
#
# #6
l2$c[c(1, 4)]
#>   a   d 
#> "a" "d"
#
# #7
l2$c[c("a", "d")]
#>   a   d 
#> "a" "d"
#
# #8
l2["c"][["c"]][c("a", "d")]
#>   a   d 
#> "a" "d"
#
# #9
l[2:3][3]  # doesn't work
#> [[1]]
#> NULL
```
In 2.1.3, please describe (e.g. class, data structure, number of elements) a and l in Chunk #4.
In 2.1.3 Chunk #7, please describe the operation that we just performed, and what objects are used in it.

CODE:
1. Create a vector a, with values 20:30, a vector b holding all letters in the alphabet.
```{r}
a <-c(20:30)
a
b<-letters
b
```
2. Assign letters as the names for vector a, such that a[1] gets the name “a”, a[2] gets named “b”, etc.
```{r}
names(a)<-letters[1:11]
a
```
3. Combine those vectors in to a list l, assigning names a and b to the two list elements
```{r}
l<-list(a,b)
names(l)<-c("a","b")
l
l$b
```
Select from a as follows:
All values >= 26
```{r}
l$a[a>=26]
```
The 1st and 7th element
```{r}
l$a[c(1,7)]
```
The last element and the second to last element (extra marks if you use length to find both index numbers)
```{r}
l$a[c(length(l$a),length(l$a)-1)]
```
Select from a the values named “a”, “c”, and “g”
```{r}
l$a[c("a","c","g")]
```
Select from l:
The first element by integer index

```{r}
l[[1]]
```
The first element by integer index, so that it returns as a 1-element list
```{r}
l[1]
```
All values in the element named a that are < than 25
```{r}
l$a[a<25]
```
All values in the element named a that are equal to 25
```{r}
l$a[a==25]
```
All values in the second element that are contained in the vector of letters c("d", "e", "f")
```{r}
l$b [b%in% c("d", "e", "f")]
```
MATRIX
```{r}
m <- cbind(1:5, 11:15, 21:25)
m
m[1,1]
m[2,2]
m[1:2,2:3]
m[c(1, 4), c(1, 3)]
m[-1, -3]
m[-c(1:4), -c(1:2)]
```
```{r}
colnames(m) <- letters[1:3]
rownames(m) <- letters[1:5]
m
```
```{r}
m["a", "a"]
```
```{r}
m[c("a", "d"), c("a", "b")]
```
```{r}
m[letters[c(1, 5)], letters[1:2]]
```
```{r}
m["a", -c("5")]
```
```{r}
m["a", -(colnames(m) == "a")]
```
```{r}
m[-which(rownames(m) %in% c("a", "b", "c")), "a"]
```
A matrix value can also be accessed with a single vector:
```{r}
m[1:5]
m[6:10]
m[1:length(m)]
```
```{r}
set.seed(1)
d <- data.frame(a = letters[1:4], b = 1:4, c = runif(n = 4, min = 0, max = 20))
d
```

```{r}
is.list(d)
```
The second major difference is that single vector indexing is applied to columns only in data.frames, not by row, column.
```{r}
d[1]

d[c("a", "c")]

d[1:3]

d[1:4]
```

```{r}
d[d$a %in% c("a", "c"), "c"]
```
```{r}
d[d$b > 2 & d$c < 18, 1:3]
```

```{r}
m <- cbind(1:5, 11:15, 21:25)
m
```


```{r}
m[1:4, 1:2] <- -9
m
```

```{r}
colnames(m)<-letters[1:3]
rownames(m)<-letters[1:5]
m[m[, "c"] == 0, 3] <- 24:25
m
```
```{r}
d
d[d$b <= 2, "a"] <- "zzz"
d
```

```{r}
ds <- d[d$b >= 2 & d$c < 18, 2:3]
d[d$b >= 2 & d$c < 18, 2:3] <- 11:14
d
```
```{r}
d[d$b >= 11 & d$c <= 14, 2:3] <- ds
d
```

```{r}
d[d["a"]=="zzz","a"]<-letters[1:2]
d
```
```{r}
d[[3]] <- 10^d$b
d
```

```{r}
set.seed(1)
d <- data.frame(a = letters[1:7], b = 1:7, c = runif(n = 7, min = 0, max = 20))
d
```

```{r}
install.packages("tidyverse")
library(dplyr)
```

```{r}
d %>% filter(a %in% c("a", "e")) %>% select(a, b)
```
```{r}
d %>% filter(c > 7 & c < 18) %>% select(-b)
```

```{r}
d %>% mutate(a = ifelse(a %in% c("a", "e"), "zzz", a))
```
```{r}
d %>% mutate(a = ifelse(a %in% c("a", "e"), "zzz", a), 
             c = ifelse(c > 7 & c < 18, -9999, c))
```
```{r}
d %>% mutate(a = ifelse(a %in% c("a", "e"), "zzz", a)) %>%  
  mutate(c = ifelse(c > 7 & c < 18, -9999, c)) %>% mutate(b = b + 10) %>% 
  mutate(d = b^2)
```
1. Create a matrix m with 10 rows and 3 columns. Make the 1st column have values 1:10, the second column 11:20, and the third 21:30
```{r}
m <- cbind(1:10, 11:20, 21:30)
m
```
2. Select rows 4 and 5 and columns 2 and 3 from the matrix
```{r}
m[c(4,5),c(2,3)]
```

3. Name the matrix columns “a”, “b”, and “c”
```{r}
colnames(m)<-(letters[1:3])
m
```

4. Select the values from column “b” that are greater than 14 and less than or equal to 18
```{r}
m[m[,"b"]>14 & m[,"b"]<=18,"b"]
```

5. Convert m to a data.frame d
```{r}
d<- as.data.frame(m)
d
```

6. Select from d$a the values >4 and replace them with the value -1
```{r}
d[d$a>4,"a"]<- -1
d
```

7. Replace the values in column c with the first 10 letter of the alphabet
```{r}
d$c<-(letters[1:10])
d
```

8. Combine m and d into a list l. Select rows 2 and 3 from column b from the element containg d of list l
```{r}
l<-list(m,d)
l

#l[[2]][2:3,2]
```
```{r}
l[[2]][2:3,"b"]
```


9. Use dplyr::filter to select the values between 14 and 18 from column b of d.


```{r}
filter(d, b > 14 & b <= 18) %>% select(b)
```

```{r}
a <- 1:10
log(a) 
```

```{r}
exp(log(a))
```
```{r}
b <- 1:5
m <- cbind(v1 = 1:5, v2 = 11:15)
# 
m
```
```{r}
m2 <- cbind(c(10, 20), c(5, 10))
#
m2
```
```{r}
d <- data.frame(m, v3 = 101:105, v4 = letters[1:nrow(m)],
                stringsAsFactors = FALSE)
d
```
```{r}
b * m 
```
```{r}
b[length(b)] * m
```
```{r}
d[1, 1:3] * m[nrow(m), 1]
```

```{r}
d$v1 * m
```
##3.1.1 Questions

1. In Chunk 19 #3 explain the order of calculations.
alternates between the first and fifth elements of b, multiplying down each column of m. 

2. In Chunk 19 #8, what would the order of calculations be if m had 3 columns?

the third column of the df would be multiplied by the 3rd row of the matrix

3. In Chunk 19 #9, describe the dimensions of the subsets extracted from d and m, and the order of calculations.

d[1, 1:3] extracts the 1st column, rows 1 through 3
m[nrow(m), 1] extracts the 5th row, column 1 (m=5)

##3.1.2 Code

1. Repeat the code in Chunk 19, but replace multiplication with division, addition, and subtraction

```{r}
# Chunk 19
b <- 1:5
m <- cbind(v1 = 1:5, v2 = 11:15)
# 
m
#>      v1 v2
#> [1,]  1 11
#> [2,]  2 12
#> [3,]  3 13
#> [4,]  4 14
#> [5,]  5 15
m2 <- cbind(c(10, 20), c(5, 10))
#
m2
#>      [,1] [,2]
#> [1,]   10    5
#> [2,]   20   10
d <- data.frame(m, v3 = 101:105, v4 = letters[1:nrow(m)],
                stringsAsFactors = FALSE)
d
#>   v1 v2  v3 v4
#> 1  1 11 101  a
#> 2  2 12 102  b
#> 3  3 13 103  c
#> 4  4 14 104  d
#> 5  5 15 105  e
#
# #1
b / m 

# #2
b[length(b)] / m

# #3
b[c(1, 5)] / m

# #4
m / m

# #5
#m / m2

# #6
m / m2[, 1]

# #7
#
#d / m

# #8
d[, 1:3] * m
#>   v1  v2  v3
#> 1  1 121 101
#> 2  4 144 204
#> 3  9 169 309
#> 4 16 196 416
#> 5 25 225 525
#
# #9
d[1, 1:3] / m[nrow(m), 1]

#10
d$v1 / m

# 11
d$v1 / m2
#> Warning in d$v1 * m2: longer object length is not a multiple of shorter object length
#> Error in eval(expr, envir, enclos): dims [product 4] do not match the length of object [5]
```

2. Find the sine and cosine of matrix m

```{r}
sin(m)
cos(m)
```


3. For the rowSums, colSums, rowMeans, and colMeans of the numeric subset of d
```{r}
d
rowSums(d[1:3])
```

```{r}
colSums(d[1:3])

rowMeans(d[1:3])

colMeans(d[1:3])
```

```{r}
sscript <- c("st", "nd", "rd", "th")  # vector of superscripts  
for(i in 1:4) {  # for loop with iterator i over vector 1:4
  stmnt <- paste0(letters[i], " is the ", i, sscript[i],
                  " letter in the alphabet")
  print(stmnt)
}
```

```{r}
sscript <- c("st", "nd", "rd", "th")  # vector of superscripts  
par(mfrow = c(1, 4), mar = c(0, 0, 1, 0.5))
for(i in 1:4) {  
  stmnt <- paste0(letters[i], " is the ", i, sscript[i],
                  " letter in the alphabet")
  plot(1:4, rep(3, 4), ylim = c(1, 5), pch = letters[1:4], axes = FALSE,
       xlab = "", ylab = "", main = stmnt, cex = 2) 
  points(i, 3, pch = letters[i], col = "red", cex = 2)
}
```

```{r}
for(i in 1:20) {
  if(i %in% seq(5, 20, by = 5)) {
    print(paste(i, "is divisible by 5"))
  } 
  if(i == 10) {
    print(paste(i, "is halfway to 20"))
  } 
  if(i == 20) {
    print(paste(i, "is the last number. Finished!"))
  }
}
```

```{r}
## using ifelse to capture values in a new vector
a <- 1:10
b <- ifelse(a < 7, 0, a)
b
```

## 4.2.1.1 Code
1. Copy Chunk 21’s code. Change the iterator vector to 1:5 and re-run the code chunk. What happens? Do you need to make any changes to make a better result?

```{r}
sscript <- c("st", "nd", "rd", "th")  # vector of superscripts  
for(i in 1:length(letters)) {  # for loop with iterator i over vector 1:4
  if(i<4){
  stmnt <- paste0(letters[i], " is the ", i, sscript[i],
                  " letter in the alphabet")
  }else{
    stmnt<-paste0(letters[i], " is the ",i,"th letter in the alphabet")
  }
  print(stmnt)  # print statement
}
```

2. Copy Chunk 22’s code.
-Comment out the lines beginning with points and run the code to see what that line does
-Delete the “axes = FALSE” part of the call to plot and see what does.

```{r}
sscript <- c("st", "nd", "rd", "th")  # vector of superscripts  
par(mfrow = c(1, 4), mar = c(0, 0, 1, 0.5))
for(i in 1:4) {  
  stmnt <- paste0(letters[i], " is the ", i, sscript[i],
                  " letter in the alphabet")
  plot(1:4, rep(3, 4), ylim = c(1, 5), pch = letters[1:4], axes = FALSE,
       xlab = "", ylab = "", main = stmnt, cex = 2) 
  points(i, 3, pch = letters[i], col = "red", cex = 2)
}
```


3. Copy Chunk 26’s code:
- Change the if statement within condition 2 such that i >= 3. Run it and inspect the result
- Now change the second half of the statement so that i <= 8. What’s the result?

```{r}
for(i in 1:10) {
  if(i < 5) {  # condition 1
    print(paste(i, "is less than", i + 1))
  } else if(i >= 3 & i <= 8) {  # condition 2
    print(paste(i, "is between", i - 1, "and", i + 1))
  } else {  # remaining conditions
    print(paste(i, "is greater than", i - 1))
  }
}
```


4. Create a for loop that iterates over a vector 1:20. Insert a condition into it such that it only prints out a result when the iterator’s value is 11

```{r}
vec<-1:20
vec
for(i in vec){
  if(i==11){
    print("11!")
  }
}
```

```{r}
dat_list <- list(data.frame(a = 1:10, b = 21:30), 
                 data.frame(a = 31:40, b = 41:50),
                 data.frame(a = 51:60, b = 61:70))
dat_list
```

```{r}
l<-lapply(dat_list, rowMeans)
l
```

```{r}
inverse_log10 <- function(x) 10^x
lapply(1:4, inverse_log10)
unlist(lapply(1:4, inverse_log10))
```

```{r}
dat_list2 <- c(dat_list, mean)  # add another element to dat_list
lapply(1:length(dat_list2), function(x) {  
  d <- dat_list2[[x]]  # extract element of list
  if(is.data.frame(d)) {  # check if it is a data.frame
    d[d == 999] <- NA  # convert any 999 values to NA 
    o <- c(colSums(d, na.rm = TRUE), # column sums, dropping NAs
           "total" = sum(d, na.rm = TRUE)) #  sum dropping NAs
  } else { # if it is not a data.frame, make an error statement
    o <- paste("Operation not valid for a", class(d))  
  }
  return(o)  # return result
})
```

```{r}
flist <- list(mean, sd, range)
lapply(1:3, function(x) flist[[x]](unlist(dat_list[[1]])))
```

```{r}
sapply(dat_list, rowSums)
```

```{r}
dat <- dat_list[[2]]
dat
```

```{r}
apply(X = dat, MARGIN = 1, FUN = sum)
```

```{r}
apply(X = dat, MARGIN = 2, FUN = sum)
```
## 4.3.5 Practice
```{r}
install.packages("swirl")
library(swirl)
swirl()

```

## 4.3.5.1 Code
1. Copy and run Chunk 23 and 29’s code to recreate dat_list. Copy Chunk 30’s code and:
- Set na.rm = FALSE in the call to colSums
- Re-run the code to check how the answer changes.
```{r}
dat_list <- list(data.frame(a = 1:10, b = 21:30), 
                 data.frame(a = 31:40, b = 41:50),
                 data.frame(a = 51:60, b = 61:70))
for(i in dat_list) print(rowSums(i))
```

```{r}
dat_modify <- function(x) {
  x[1:3, 1] <- 999
  return(x)
}
dat_list <- lapply(dat_list, dat_modify)
dat_list
```

```{r}
dat_list2 <- c(dat_list, mean)  # add another element to dat_list
dat_list2

```
```{r}
lapply(1:length(dat_list2), function(x) {  
  d <- dat_list2[[x]]  # extract element of list
  if(is.data.frame(d)) {  # check if it is a data.frame
    d[d == 999] <- NA  # convert any 999 values to NA 
    o <- c(colSums(d, na.rm = FALSE), # column sums, dropping NAs
           "total" = sum(d, na.rm = TRUE)) #  sum dropping NAs
  } else { # if it is not a data.frame, make an error statement
    o <- paste("Operation not valid for a", class(d))  
  }
  return(o)  # return result
}) # return result
```


2. Copy Chunk 31’s code. Remove the unlist() around dat_list[[1]]. Re-run the code and see what happens.
```{r}
flist <- list(mean, sd, range)
lapply(1:3, function(x) flist[[x]](dat_list[[1]]))
```

3. Write two lapply statements to apply the functions mean and sd to dat_list
```{r}
dat_list
sapply(dat_list, mean)
```

4. Do the same for sapply
```{r}

```

5. Write an lapply with an anonymous function that iterates through dat_list, selecting the 2nd column, rows 1 and 2 from each data.frame, and then sum that subset.
```{r}
lapply(dat_list, function(x) sum(x[c(1, 2), 2]))
```

